---
title: "Evoman Specialist Statistical tests"
author: "Meena Kerolos"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
header-includes:
  - \usepackage{leading}
  - \leading{9pt}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=3)
```


```{r read_data, include=FALSE}
# Read the data
results = read.table(file="../results/Specialist_final_2/results_test.csv",header=TRUE, sep=";")
results$method = as.factor(results$method)
results$enemy = as.factor(results$enemy)
```


```{r load_libraries, include=FALSE}
# Load libraries
if(!require(rcompanion)){install.packages("rcompanion")}
if(!require(FSA)){install.packages("FSA")}
library(rcompanion)
```

# Using Scheirer–Ray–Hare test

Scheirer–Ray–Hare test is a non-parametric test (doesn't have assumptions on data).

```{r}
scheirerRayHare(gain ~ method + enemy, data = results)
```

There is indeed an interaction between method and enemy
The enemy is already known to have an effect on the outcome
We can't test the actual effect of the method because it will already
have an effect through the interaction.


# Separate experiments

Let's treat each enemy as a separate experimental setup
The comparison is made between methods and we are interested in testing
the difference between the populations of the set of outputs coming from each
method.
For that we will use Mann-Whitney test

## Enemy 1

```{r}
enemy1_Static = results[(results$enemy == 1) & (results$method == 'EvomanAnnNeatStatic'),]$gain
enemy1_Dynamic = results[(results$enemy == 1) & (results$method == 'EvomanAnnNeatDynamic'),]$gain
z = wilcox.test(enemy1_Static, enemy1_Dynamic); z
p_value = z$p.value
```

This test is testing the null hypothesis H0: F = G which means that both sets are coming from the same population.

With a p-value of `r p_value` < 0.05, we can reject that hypothesis and conclude that there is indeed a treatment effect. Here are the mean and median of each set as an estimators of the location of each population:

- gain for Static method: mean=`r mean(enemy1_Static)`, median=`r median(enemy1_Static)`
- gain for Dynamic method: mean=`r mean(enemy1_Dynamic)`, median=`r median(enemy1_Dynamic)`

Considering the mean of each set as an estimator to the location of the population we see that the Static method results in more gain than the Dynamic method for enemy 1.


## Enemy 2

```{r}
enemy2_Static = results[(results$enemy == 2) & (results$method == 'EvomanAnnNeatStatic'),]$gain
enemy2_Dynamic = results[(results$enemy == 2) & (results$method == 'EvomanAnnNeatDynamic'),]$gain
z = wilcox.test(enemy2_Static, enemy2_Dynamic); z
p_value = z$p.value
```

This test is testing the null hypothesis H0: F = G which means that both sets are coming from the same population.

With a p-value of `r p_value` < 0.05, we can reject that hypothesis and conclude that there is indeed a treatment effect. Here are the mean and median of each set as an estimators of the location of each population:

- gain for Static method: mean=`r mean(enemy2_Static)`, median=`r median(enemy2_Static)`
- gain for Dynamic method: mean=`r mean(enemy2_Dynamic)`, median=`r median(enemy2_Dynamic)`

Considering the mean of each set as an estimator to the location of the population we see that the Static method results in more gain than the Dynamic method for enemy 2.

# Enemy 3
```{r}
enemy3_Static = results[(results$enemy == 3) & (results$method == 'EvomanAnnNeatStatic'),]$gain
enemy3_Dynamic = results[(results$enemy == 3) & (results$method == 'EvomanAnnNeatDynamic'),]$gain
z = wilcox.test(enemy3_Static, enemy3_Dynamic); z
p_value = z$p.value
```

This test is testing the null hypothesis H0: F = G which means that both sets are coming from the same population.

With a p-value of `r p_value` < 0.05, we can reject that hypothesis and conclude that there is indeed a treatment effect. Here are the mean and median of each set as an estimators of the location of each population:

- gain for Static method: mean=`r mean(enemy3_Static)`, median=`r median(enemy3_Static)`
- gain for Dynamic method: mean=`r mean(enemy3_Dynamic)`, median=`r median(enemy3_Dynamic)`

Considering the mean of each set as an estimator to the location of the population we see that the Static method results in more gain than the Dynamic method for enemy 3.


# Enemy 6
```{r}
enemy6_Static = results[(results$enemy == 6) & (results$method == 'EvomanAnnNeatStatic'),]$gain
enemy6_Dynamic = results[(results$enemy == 6) & (results$method == 'EvomanAnnNeatDynamic'),]$gain
z = wilcox.test(enemy6_Static, enemy6_Dynamic); z
p_value = z$p.value
```

This test is testing the null hypothesis H0: F = G which means that both sets are coming from the same population.

With a p-value of `r p_value` > 0.05, we CANNOT reject the null hypothesis and conclude that both methods doesn't have a different effect of the gain. Here are the mean and median of each set as an estimators of the location of each population:

- gain for Static method: mean=`r mean(enemy6_Static)`, median=`r median(enemy6_Static)`
- gain for Dynamic method: mean=`r mean(enemy6_Dynamic)`, median=`r median(enemy6_Dynamic)`

# Enemy 7
```{r}
enemy7_Static = results[(results$enemy == 7) & (results$method == 'EvomanAnnNeatStatic'),]$gain
enemy7_Dynamic = results[(results$enemy == 7) & (results$method == 'EvomanAnnNeatDynamic'),]$gain
z = wilcox.test(enemy7_Static, enemy7_Dynamic); z
p_value = z$p.value
```

This test is testing the null hypothesis H0: F = G which means that both sets are coming from the same population.

With a p-value of `r p_value` < 0.05, we can reject that hypothesis and conclude that there is indeed a treatment effect. Here are the mean and median of each set as an estimators of the location of each population:

- gain for Static method: mean=`r mean(enemy7_Static)`, median=`r median(enemy7_Static)`
- gain for Dynamic method: mean=`r mean(enemy7_Dynamic)`, median=`r median(enemy7_Dynamic)`

Considering the mean of each set as an estimator to the location of the population we see that the Dynamic method results in more gain than the Static method for enemy 7.


