---
title: "Evoman Generalist Statistical tests"
author: "Meena Kerolos"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
header-includes:
  - \usepackage{leading}
  - \leading{9pt}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=3)
```


```{r read_data, include=FALSE}
# Read the data
results = read.table(file="../results/generalist_final_1/test_all_enemies_results.csv",header=TRUE, sep=",")
results$method = as.factor(results$method)
results$enemy = as.factor(results$enemy)
results$enemy_group = as.factor(results$enemy_group)
results$gain = results$player_life - results$enemy_life
```


```{r load_libraries, include=FALSE}
# Load libraries
if(!require(rcompanion)){install.packages("rcompanion")}
if(!require(FSA)){install.packages("FSA")}
library(rcompanion)
```


# Separate experiments

We will consider training against enemy group `7_8` as a different experimental setup
from training against enemy group `3_7_8`. Therefore, training against one enemy group
contains a set of experiments to compare between two EA instances. Each EA instance
is tested against each of the 8 enemies. The experimental setup is factorial block
design where the method is considered the treatment factor, the enemy is considered
the block factor and the gain is the outcome.

Scheirer–Ray–Hare test will be used as it is a non-parametric test (doesn't have
assumptions on data).

# Enemy Group 3_7_8

```{r}
results_3_7_8 = results[(results$enemy_group == '3_7_8'),]
z = scheirerRayHare(gain ~ method + enemy, data = results_3_7_8); z
```

With a p-value of `r z$p.value[3]` > 0.05, there is no significant interaction between
method and enemy and we can look at the effect of each of them. It is known that
enemy effect is significant and p-value of `r z$p.value[2]` < 0.05 for the enemy
effect supports it. However, a p-value of `r z$p.value[1]` > 0.05 for the method
effect is not significant. Therefore, GA and DE doesn't give a significantly different
gain when trained on enemy group `3_7_8` and tested against all enemies.


# Enemy Group 7_8

```{r}
results_7_8 = results[(results$enemy_group == '7_8'),]
z = scheirerRayHare(gain ~ method + enemy, data = results_7_8); z
```

With a p-value of `r z$p.value[3]` > 0.05, there is no significant interaction between
method and enemy and we can look at the effect of each of them. It is known that
enemy effect is significant and p-value of `r z$p.value[2]` < 0.05 for the enemy
effect supports it. A p-value of `r z$p.value[1]` < 0.05 for the method
effect is significant. Therefore, GA gives a higher gain than DE when trained against
enemy group `7_8` and tested against all enemies.

Here are the mean and median estimators of the gain for each method:

```{r echo=F}
results_7_8_GA_gain = results_7_8[(results_7_8$method == 'EvomanGeneralist_GA'),]$gain
results_7_8_DE_gain = results_7_8[(results_7_8$method == 'EvomanGeneralist_DE'),]$gain
```

- gain for GA method trained against enemy group `7_8`: mean=`r mean(results_7_8_GA_gain)`, median=`r median(results_7_8_GA_gain)`
- gain for DE method trained against enemy group `7_8`: mean=`r mean(results_7_8_DE_gain)`, median=`r median(results_7_8_DE_gain)`


